PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$

### 2014-09-08 qfenv_paper/qf_environment.tex
Some changes to section 3

\title{PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 1$}
\author{
ChangHoon~Hahn\altaffilmark{1}, 
Michael~Blanton\altaffilmark{1}, 
Alison~Coil\altaffilmark{2}, 
Richard~Cool\altaffilmark{3}, 
Daniel Eisenstein\altaffilmark{4},
John Moustakas\altaffilmark{5}, 
Ken Wong\altaffilmark{6}, 
Guangtun Zhu\altaffilmark{7}
}
\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003}
\altaffiltext{2}{Center for Astrophysics and Space Sciences, Department of Physics, University of California, 9500 Gilman Dr., La Jolla, CA 92093}
\altaffiltext{3}{MMT Observatory, University of Arizona, 1540 E Second Street, Tucson AZ 85721}
\altaffiltext{4}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{5}{Department of Physics and Astronomy, Siena College, 515 Loudon Road, Loudonville, NY 12211}
\altaffiltext{6}{Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721} 
\altaffiltext{7}{Department of Physics and Astronomy, The Johns Hopkins University, 3400 North Charles Street, Baltimore, MD 21218} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We investigate the effects of galaxy environment on the evolution of the quiescent fraction ($f_{\rm{Q}}$) from $z =0.8 $ to $ 0.0$ using spectroscopic redshifts and multi-wavelength imaging data from the PRism MUlti-object Survey (PRIMUS) and the Sloan Digitial Sky Survey (SDSS). Our stellar mass limited galaxy sample consists of $\sim 13,000$ PRIMUS galaxies within $z = 0.2-0.8$ and $\sim 64,000$ SDSS galaxies within $z = 0.05-0.12$. We classify the galaxies as quiescent or star-forming based on an evolving specific star formation cut, and as low or high density environments based on fixed cylindrical aperture environment measurements on a volume-limited Environment Definition Population (from PRIMUS and SDSS). For quiescent and star-forming galaxies in low or high density environments, we examine the evolution of their stellar mass function (SMF) evolution. Then using the SMFs we compute the $f_{\rm{Q}}(\mathcal{M}_{*})$ and quantify the evolution of $f_{\rm{Q}}(\mathcal{M}_{*})$ within our redshift range. We find that in both low and high density environments the quiescent fraction increases with cosmic time. Moreover, we illustrate that galaxy populations in high density environments have significantly higher quiescent fractions than galaxy populations in low density environments. Furthermore, by comparing the quiescent fraction evolutions, we find that the quiescent fraction evolves noticeably more for galaxy populations in high density environments. These results suggest that quenching due to the environment has a moderate effect on the quiescent fraction evolution and provide constraints on quenching mechanisms in high density environments such as groups and clusters. 
\end{abstract}

@@

In comparison to \cite{bundy06a} and \cite{Bolzonella:2010aa}, some of the noted trends are not immediately evident. However, this is mostly due to the conflicting mass limits proposed in each study. Not to mention, only part of the redshift range probed by \cite{bundy06a} overlap with our results. Overall, we find that the trends we note in the SMF evolution above are in good agreement with results from DEEP2 and zCOSMOS.

Observing the evolutionary trends in SMF for each of these sub-populations simultaneously provides a naive narrative of the different galaxy evolutionary tracks involving environment and quenching. For example, the decrease in the massive star-forming galaxies in low density environment over cosmic time can be attributed to the transition of those galaxies to any of the other panels. The star-forming galaxies in low density environments that have quenched their star formation over time are possibly responsible for the increase of the quiescent, low density environment SMF over time. The star-forming galaxies that fall into higher density environments explain the increase in the star-forming high density environment SMF below the knee. Finally, star-forming galaxies in high density environments that have quenched their star-formation, quiescent galaxies that have transitioned from low to high density environments, and star-forming galaxies in low density environments that quench their star-formation while infalling to high density environments all contribute to the overall increase of the high environment quiescent SMF.

@@

While our mass complete subsample coupled with robust environment measurement allows us to compare SMF evolution for each of our subsamples over $z < 1.0$, we caution readers and reserve detailed analysis of the SMFs to future work due to the improvements in photometry that significantly alter the SDSS SMFs mentioned in Section \ref{sec:sdss}.

\begin{figure*}
    \begin{center}
        \leavevmode
        \epsfig{file=fig_qf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=90}
        \caption{Evolution of the quiescent fraction $f_{\rm{Q}}$ for galaxies in low (left) and high (rights) density environments for $z < 0.8$. $f_{\rm{Q}}$s were calculated using the SMFs in Figure\ref{fig:smf}, as described in text. Darker shading indicates lower redshift and the width represents the standard jackknife uncertainty.}         \label{fig:qf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUIESCENT FRACTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiescent Fraction} \label{sec:qf_const}
The SMFs calculated in the previous section illustrate the stellar mass distribution of our galaxy population and its evolution over comic time. In this section, using the SMFs of our subsamples, we compare the quiescent and the star-forming populations by calculating the fraction of galaxies that have quenched their star-formation, the quiescent fraction.

With the sufficient statistics available from SDSS and PRIMUS, we evaluate the quiescent fraction in bins of stellar mass and redshift for low and high density environments (Section \ref{sec:qfevol}). More specifically, by analyzing the quiescent fraction with respect to mass, redshift, and environment we are able to compare the quiescent fraction evolution in low and high density environments. Through this comparison, in Section \ref{sec:env_qf_evol}, we reveal the subtle environmental effects in quenching star formation that are often obscured by the underlying correlations among galactic properties such as color-mass and mass-density relations (\citealt{Cooper:2010aa}). Furthermore by quantifying the environmental effects of star-formation quenching, we are able to provide constraints on the various proposed environmental quenching mechanisms, such as strangulation and ram-pressure stripping (\citealt{McCarthy:2008aa}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EVOLUTION OF THE QUIESCENT FRACTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolution of the Quiescent Fraction} \label{sec:qfevol}
From the SMF number densities ($\Phi$) computed in the previous section, the quiescent fraction is computed as follows, 
\begin{equation}
f_{\rm{Q}} = \frac{\Phi_{Q}}{\Phi_{SF}+\Phi_{Q}}.
\end{equation}
$\Phi_{Q}$ and $\Phi_{SF}$ are the total number of galaxies per unit volume in stellar mass bin of $\Delta(\rm{log} \: \mathcal{M}) = 0.20 \: {\it dex}$ for the quiescent and star-forming subsamples, respectively (Equation \ref{eq:phi}). We compute $f_{\rm{Q}}$ for high and low density environments for all redshift bins as plotted in Figure \ref{fig:qf}, which shows the evolution of $f_{\rm{Q}}$ for high (right panel) and low (left panel) density environments. As in Figure \ref{fig:smf}, the evolution of the quiescent fraction over cosmic time is represented in the shading (darker with lower redshift) and the uncertainty is represented by the width. For the uncertainty in the quiescent fraction, we use the standard jackknife technique, following the same steps as the SMF uncertainty in Section \ref{sec:smfcalc}.

>>>>

\title{PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at $z < 0.8$}
\author{
ChangHoon~Hahn\altaffilmark{1}, 
Michael~Blanton\altaffilmark{1}, 
Alison~Coil\altaffilmark{2}, 
Richard~Cool\altaffilmark{3}, 
Daniel Eisenstein\altaffilmark{4},
John Moustakas\altaffilmark{5}, 
Ken Wong\altaffilmark{6}, 
Guangtun Zhu\altaffilmark{7}
}
\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003}
\altaffiltext{2}{Center for Astrophysics and Space Sciences, Department of Physics, University of California, 9500 Gilman Dr., La Jolla, CA 92093}
\altaffiltext{3}{MMT Observatory, University of Arizona, 1540 E Second Street, Tucson AZ 85721}
\altaffiltext{4}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{5}{Department of Physics and Astronomy, Siena College, 515 Loudon Road, Loudonville, NY 12211}
\altaffiltext{6}{Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721} 
\altaffiltext{7}{Department of Physics and Astronomy, The Johns Hopkins University, 3400 North Charles Street, Baltimore, MD 21218} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We investigate the effects of galaxy environment on the evolution of the quiescent fraction ($f_{\rm{Q}}$) from $z =0.8 $ to $ 0.0$ using spectroscopic redshifts and multi-wavelength imaging data from the PRism MUlti-object Survey (PRIMUS) and the Sloan Digitial Sky Survey (SDSS). Our stellar mass limited galaxy sample consists of $\sim 13,000$ PRIMUS galaxies within $z = 0.2-0.8$ and $\sim 64,000$ SDSS galaxies within $z = 0.05-0.12$. We classify the galaxies as quiescent or star-forming based on an evolving specific star formation cut, and as low or high density environments based on fixed cylindrical aperture environment measurements on a volume-limited Environment Definition Population (from PRIMUS and SDSS). For quiescent and star-forming galaxies in low or high density environments, we examine the evolution of their stellar mass function (SMF) evolution. Then using the SMFs we compute the $f_{\rm{Q}}(\mathcal{M}_{*})$ and quantify the evolution of $f_{\rm{Q}}(\mathcal{M}_{*})$ within our redshift range. We find that in both low and high density environments the quiescent fraction increases with cosmic time. Moreover, we illustrate that galaxy populations in high density environments have significantly higher quiescent fractions than galaxy populations in low density environments. Furthermore, by comparing the quiescent fraction evolutions, we find that the quiescent fraction evolves noticeably more for galaxy populations in high density environments. These results suggest that quenching due to the environment has a moderate effect on the quiescent fraction evolution and provide constraints on quenching mechanisms in high density environments such as groups and clusters. 
\end{abstract}

@@

In comparison to results from DEEP2 (\citealt{bundy06a}) and zCOSMOS (\citealt{Bolzonella:2010aa}), some of the noted trends are not immediately evident. However, this is mostly due to the conflicting mass limits proposed in each study. In addition, only part of the redshift range probed by \cite{bundy06a} overlap with our results. Overall, we find that the trends we note in the SMF evolution above are in good agreement with \cite{bundy06a} and \cite{Bolzonella:2010aa}.

Observing the evolutionary trends in SMF for each of these sub-populations provides a narrative of the different galaxy evolutionary tracks involving environment and quenching. For example, the decrease in the massive star-forming galaxies in low density environments over cosmic time can be attributed to the transition of those galaxies to any of the other panels. The star-forming galaxies in low density environments that have quenched their star formation over time are possibly responsible for the increase of the quiescent, low density environment SMF over time. The star-forming galaxies that fall into higher density environments explain the increase in the star-forming high density environment SMF below the knee. Finally, star-forming galaxies in high density environments that have quenched their star-formation, quiescent galaxies that have transitioned from low to high density environments, and star-forming galaxies in low density environments that quench their star-formation while infalling to high density environments all contribute to the overall increase of the high environment quiescent SMF.

@@

While our mass complete subsample coupled with robust environment measurement allows us to compare SMF evolution for each of our subsamples out to $z=0.8$, we caution readers and reserve detailed analysis of the SMFs to future work due to the improvements in photometry that significantly alter the SDSS SMFs mentioned in Section \ref{sec:sdss}.

\begin{figure*}
    \begin{center}
        \leavevmode
        \epsfig{file=fig_qf_cylr2h25_thresh75_bin0_20_sdss_z0_05_0_12_primus_z0_2_1_0_lowenv0_5_highenv3_0_lit_primuszerr.eps, height=0.75\textwidth,angle=90}
        \caption{Evolution of the quiescent fraction $f_{\rm{Q}}$ for galaxies in low (left) and high (rights) density environments for $z < 0.8$. $f_{\rm{Q}}$s were calculated using the SMFs in Figure\ref{fig:smf}, as described in text. Darker shading indicates lower redshift and the width represents the standard jackknife uncertainty.}         \label{fig:qf}
    \end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% QUIESCENT FRACTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quiescent Fraction} \label{sec:qf_const}
The SMFs calculated in the previous section illustrate the stellar mass distribution of our galaxy population and its evolution over cosmic time. In this section, using the SMFs of our subsamples, we compare the quiescent and the star-forming populations by calculating the fraction of galaxies that have quenched their star-formation, the quiescent fraction.

With the sufficient statistics available from SDSS and PRIMUS, we evaluate the quiescent fraction in bins of stellar mass and redshift for low and high density environments (Section \ref{sec:qfevol}). More specifically, by analyzing the quiescent fraction with respect to mass, redshift, and environment we are able to compare the quiescent fraction evolution in low and high density environments. Through this comparison, in Section \ref{sec:env_qf_evol}, we reveal the subtle environmental effects in quenching star formation that are often obscured by the underlying correlations among galactic properties such as color-mass and mass-density relations (\citealt{Cooper:2010aa}). Furthermore by quantifying the environmental effects of star-formation quenching, we are able to provide constraints on the various proposed environmental quenching mechanisms, such as strangulation and ram-pressure stripping (\citealt{McCarthy:2008aa}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% EVOLUTION OF THE QUIESCENT FRACTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolution of the Quiescent Fraction} \label{sec:qfevol}
From the SMF number densities ($\Phi$) computed in the previous section, the quiescent fraction is computed as follows, 
\begin{equation}
f_{\rm{Q}} = \frac{\Phi_{Q}}{\Phi_{SF}+\Phi_{Q}}.
\end{equation}
$\Phi_{Q}$ and $\Phi_{SF}$ are the total number of galaxies per unit volume in stellar mass bin of $\Delta(\rm{log} \: \mathcal{M}) = 0.20 \: dex$ for the quiescent and star-forming subsamples, respectively (Equation \ref{eq:phi}). We compute $f_{\rm{Q}}$ for high and low density environments for all redshift bins as plotted in Figure \ref{fig:qf}, which shows the evolution of $f_{\rm{Q}}$ for high (right panel) and low (left panel) density environments. As in Figure \ref{fig:smf}, the evolution of the quiescent fraction over cosmic time is represented in the shading (darker with lower redshift) and the uncertainty is represented by the width. For the uncertainty in the quiescent fraction, we use the standard jackknife technique, following the same steps as the SMF uncertainty in Section \ref{sec:smfcalc}.
